---
title: "Practical Machine Learning Project"
author: "ShuYi"
output: html_document
---

# Project Overview

The collected data is about personal activity from devices such as Jawbone Up, Nike FuelBand, and Fitbit which is part of the quantified self-movement of a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 


The goal of this project is based on the data, perform prediction on the manner in which they did the exercise. This is the "classe" variable in the training set. Any variables can be used in the prediction and generate a report that describe how the model was build and how cross validation execute and what would be the possible sample error and why the decision is made. 


Use prediction model to predict 20 difference test cases. 



# Data Processing

#### Load Library
```{r, message=FALSE}
library(caret)
library(randomForest)
library(rpart)
```


#### Data Cleaning
```{r}
# Training Data
TrainingData <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))

# Eliminate blank & Missing Value
TrainingDataSet <- TrainingData[,colSums(is.na(TrainingData)) == 0]

# Exclude not required variables
TrainingDataSet   <-TrainingDataSet[,-c(1:7)]
```

```{r}
# Testing Data
TestingData <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))

# Eliminate blank & Missing Value
TestingDataSet <- TestingData[,colSums(is.na(TestingData)) == 0]

# Exclude not required variables
TestingDataSet <-TestingDataSet[,-c(1:7)]

```

#### Data Partition
```{r}
# Partition 60% of processed training dataset
PmlDataPartition <- createDataPartition(y=TrainingDataSet$classe, p=0.6, list=FALSE)
partDataTrain <- TrainingDataSet[PmlDataPartition, ]
partDataTest <- TrainingDataSet[-PmlDataPartition, ]

# View classe levels by frequency
plot(partDataTrain$classe, col="yellow", main="Classe Levels by Frequency", xlab="classe levels", ylab="Frequency")
```

# Model Comparison and Selection
```{r}
# Tree Model
TreeModel <- train(classe ~., method="rpart", data=partDataTrain)

confusionMatrix(partDataTrain$classe, predict(TreeModel, partDataTrain))
```

```{r cache=TRUE}
# Random Forest Model
RandForestModel <- randomForest(classe ~. , data=partDataTrain, method="class")
RandForestModelVal <- predict(RandForestModel, partDataTest, type = "class")

confusionMatrix(RandForestModelVal, partDataTest$classe)
```

From the observation above, it seem like random forest model produce more accurate result. So, forest model is choosen for the prediction of the classe value of the 6 participants as this model has the 99.45% accuracy rate. The expected out-of-sample is calculated as 1 - prediction accuracy made against cross validation and the result is 0.55%.

# Prediction Result
```{r}
# Predict testing data set for the 20 test cases.
PredictResult <- predict(RandForestModel, TestingDataSet, type="class")
PredictResult

# Generate result to file
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(PredictResult)
```

# Reference
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

